---
title: "Benchmarking"
author: "Stephen Kelly"
date: "3/3/2016"
---
PWD: $HOME/projects/Benchmarking
```{r,engine='bash'}
# write a script to track time for completion of HOMER with different thread numbers

Input_BED="$HOME/projects/Benchmarking/data/Nkx2.bed"
OutDir="$HOME/projects/Benchmarking/output/timeTest_HOMERmotif"
mkdir -p "$OutDir"
Timer_Script="$HOME/projects/Benchmarking/code/time_track_HOMER.sh"
chmod +x "$Timer_Script"

qsub -q all.q -wd $OutDir -o :${OutDir}/ -e :${OutDir}/ -pe threaded 32 "$Timer_Script" "${OutDir}" "${Input_BED}" 

```

this script doesn't really work well, its using 32 slots the whole time. Need to split things up

- qsub single slot wrapper script that runs 32..1 forloop, submits qsub job to run the motif analsis
- motif analysis script that runs motif analysis 
- dedicated dir for each analysis job
- 



```{r,engine='bash'}
# write a script to track time for completion of HOMER with different thread numbers

Input_BED="$HOME/projects/Benchmarking/data/Nkx2.bed"
tmp_SAMPLE=$(basename $Input_BED)
tmp_SAMPLE=${tmp_SAMPLE%.*}

OutDir="$HOME/projects/Benchmarking/output/timeTest_HOMERmotif"
mkdir -p "$OutDir"

Timer_Track_Wrapper="$HOME/projects/Benchmarking/code/time_track_wrapper.sh"
chmod +x "$Timer_Track_Wrapper"
Timer_Script="$HOME/projects/Benchmarking/code/time_track_HOMER2.sh"
chmod +x "$Timer_Script"

# make the tmp dirs
for i in {32..1}; do
  if [ -d "$OutDir/${i}" ]; then
    rm -rf "$OutDir/${i}" # delete the dir if it exists
  fi
 mkdir -p "$OutDir/${i}"
done

# submit to wrapper 
# tmp_LOG_FILE="${OutDir}/TIMER_LOG_$(date -u +%Y%m%dt%H%M%S).txt"
tmp_LOG_FILE="${OutDir}/TIMER_LOG.txt"
qsub -q all.q -wd $OutDir -o :${OutDir}/ -e :${OutDir}/ -pe threaded 1 "$Timer_Track_Wrapper" "${OutDir}" "${Input_BED}" "$Timer_Script" "$tmp_LOG_FILE"

```

disk IO is probably a bottleneck, producing inconsistent results



```{r,engine='bash'}
# write a script to track time for completion of HOMER with different thread numbers
# don't use a wrapper script for this

Input_BED="$HOME/projects/Benchmarking/data/Nkx2_ChIPSeq_mm9.bed"
tmp_SAMPLE=$(basename $Input_BED)
tmp_SAMPLE=${tmp_SAMPLE%.*}

OutDir="$HOME/projects/Benchmarking/output/HOMERmotif"
mkdir -p "$OutDir"


tmp_Script="$HOME/projects/Benchmarking/code/time_track_HOMER3.sh"
chmod +x "$tmp_Script"

# make the tmp dirs
for i in {32..1}; do
  if [ -d "$OutDir/${i}" ]; then
    rm -rf "$OutDir/${i}" # delete the dir if it exists
  fi
 mkdir -p "$OutDir/${i}"
done

# submit to wrapper 
# tmp_LOG_FILE="${OutDir}/TIMER_LOG_$(date -u +%Y%m%dt%H%M%S).txt"
tmp_LOG_FILE="${OutDir}/TIMER_LOG_HOMER3.txt"
qsub -q all.q -wd $OutDir -o :${OutDir}/ -e :${OutDir}/ -pe threaded 32 "$tmp_Script" "${OutDir}" "${Input_BED}" "$tmp_LOG_FILE"

```



```{r,engine='bash'}

echo "stuff"
TestObj="some entry here"

R --vanilla --no-save << EOF
# myRcommands(output="results.txt")
# args <- commandArgs(TRUE)
# 
# cat("R: dput args:") # for development & troubleshooting
# cat("\n")
# cat(dput(args))
# cat("\n\n")
# sessionInfo()
cat("\nThis Is R here\n")

EOF


ls



# notes

for i in {32..1}; do
  start_time=$(date +%s)
  echo -ne "$i\t" && echo run time is $(expr $(date +%s) - $start_time) s
done


start_time=`date +%s`
<command-to-execute>
end_time=`date +%s`
echo execution time was `expr $end_time - $start_time` s.

Or in bash:

start_time=`date +%s`
<command-to-execute> && echo run time is $(expr `date +%s` - $start_time) s
```
