---
title: "Survival Analysis"
author: "Stephen Kelly"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    keep_md: yes
    number_sections: yes
    toc: yes
---

Example survival analysis workflow. Comparing the survival of patients with different mutational states in various genes; BRCA, EGFR, SOX9. 

(Simulated synthetic data shown)

```{r setup, include=TRUE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=10)
knitr::opts_chunk$set(fig.height=10)
suppressPackageStartupMessages(library("ggplot2"))
suppressPackageStartupMessages(library("FSA"))
suppressPackageStartupMessages(library("knitr"))
suppressPackageStartupMessages(library("gplots"))
suppressPackageStartupMessages(library("corrplot"))
suppressPackageStartupMessages(library("survival"))
suppressPackageStartupMessages(library("survminer"))

```

Load data and define a custom mutational state "`BRCA_EGFR_State`" based on the presence of mutations in BRCA, EGFR, and SOX9.

```{r}
# load data
surv_dat <- read.delim(file = "survival_data.tsv", header = TRUE, sep = '\t')

# add "BRCA vs. EGFR" state factor level for comparison
# - BRCA_MUT: BRCA 1, EGFR & SOX9 0
# - EGFR_MUT: EGFR 1, BRCA & SOX9 0
# 0: no mutation
# 1: mutation
surv_dat[["BRCA_EGFR_State"]] <- NA
for(i in seq(nrow(surv_dat))){
    BRCA <- surv_dat[i, "BRCA"]
    EGFR <- surv_dat[i, "EGFR"]
    SOX9 <- surv_dat[i, "SOX9"]
    if(BRCA == 1 & EGFR == 0 & SOX9 == 0){
        surv_dat[i, "BRCA_EGFR_State"] <- "BRCA_MUT"
    } else if(EGFR == 1 & SOX9 == 0 & BRCA == 0 ){
        surv_dat[i, "BRCA_EGFR_State"] <- "EGFR_MUT"
    } else if(EGFR == 1 & SOX9 == 1 & BRCA == 0 ){
        surv_dat[i, "BRCA_EGFR_State"] <- "EGFR_SOX9_MUT"
    } else {
        surv_dat[i, "BRCA_EGFR_State"] <- NA
    }
}

# convert to factor level
surv_dat[["BRCA_EGFR_State"]] <- factor(x = surv_dat[["BRCA_EGFR_State"]], levels = unique(surv_dat[["BRCA_EGFR_State"]]))
surv_dat[["Response_to_Treatment"]] <- factor(x = surv_dat[["Response_to_Treatment"]], levels = unique(surv_dat[["Response_to_Treatment"]]))
surv_dat[["Toxicity"]] <- factor(x = surv_dat[["Toxicity"]], levels = unique(surv_dat[["Toxicity"]]))
surv_dat[["Gender"]] <- factor(x = surv_dat[["Gender"]], levels = unique(surv_dat[["Gender"]]))
surv_dat[["Baseline_LDH_at_Treatment_Initiation"]] <- factor(x = surv_dat[["Baseline_LDH_at_Treatment_Initiation"]], levels = unique(surv_dat[["Baseline_LDH_at_Treatment_Initiation"]]))
surv_dat[["Stage"]] <- factor(x = surv_dat[["Stage"]], levels = unique(surv_dat[["Stage"]]))
surv_dat[["Treatment_Type"]] <- factor(x = surv_dat[["Treatment_Type"]], levels = unique(surv_dat[["Treatment_Type"]]))

```

# Age

Compare age distribution between groups and check for statistical significance.

```{r}
ggplot(data = surv_dat[which(!is.na(surv_dat[["BRCA_EGFR_State"]])), ], 
       aes(x = BRCA_EGFR_State, y = Age_at_Initiation, fill = BRCA_EGFR_State)) + 
    geom_boxplot() +
    ggtitle("Age") +
    theme_bw() +
    guides(fill=FALSE) +
    theme(panel.grid.minor = element_blank())

kruskal.test(Age_at_Initiation ~ BRCA_EGFR_State, 
             data = surv_dat[which(!is.na(surv_dat[["BRCA_EGFR_State"]])), ])

# only works for >2 levels in BRCA_EGFR_State
dunnTest(Age_at_Initiation ~ BRCA_EGFR_State, method = "by",
         data = surv_dat[which(!is.na(surv_dat[["BRCA_EGFR_State"]])), ])

```

# ANOVA

Check if there is statistical difference between group

```{r}
fit <- aov(TMB ~ BRCA_EGFR_State, data = surv_dat) 
print(fit)
summary(fit)
TukeyHSD(fit)
```

# Kruskal-Wallis

Check if there is statistical difference between group using non-parametric test

```{r}
kruskal.test(TMB ~ BRCA_EGFR_State, data = surv_dat)

# only works for >2 levels in BRCA_EGFR_State
dunnTest(TMB ~ BRCA_EGFR_State, data = surv_dat, method = "by")

```


# Response to Immunotherapy Treatment

Test for differences in response to treatment. 

- `Response_to_Treatment`: 1 = Responder, 2 = non-responder; exclude NA, '9'
- `Response_to_Treatment` vs `BRCA_EGFR_State`

### Chi Square

```{r}
# build contingency table
treatment_table <- table(surv_dat[, c('Response_to_Treatment', 'BRCA_EGFR_State')])
print(treatment_table)

balloonplot(treatment_table, 
            main ="Response to Immunotherapy Treatment\nContingency Table", 
            xlab ="Responder (1) \nvs.\nNon-Responder (2)", 
            ylab="Mutation Status"
            )

fit <- chisq.test(treatment_table)
fit

# observed counts
fit[["observed"]] 

# counts under the null
kable(round(fit[["expected"]], digits = 2))

# standardized residuals
kable(round(fit[["stdres"]] , digits = 2))

# Pearson residuals
kable(round(fit[["residuals"]] , digits = 2))

# visualization of the residuals to interpret the association between rows and columns
# # For a given cell, the size of the circle is proportional to the amount of the cell contribution.
# # Positive residuals are in blue. Positive values in cells specify an attraction (positive association) between the corresponding row and column variables.
# # Negative residuals are in red. This implies a repulsion (negative association) between the corresponding row and column variables.
corrplot(corr = fit[["residuals"]], is.cor = FALSE)
```

### Fisher Exact Test

```{r}
fisher_fit <- fisher.test(treatment_table)
fisher_fit
```

# Overall Survival

Track overall survival amount patients based on mutational state using Kaplan Meier survival curve.


```{r}
# 0=alive, 1=dead
# BRCA_EGFR_State 0=Wild Type, 1=Mutant
# 'event' must be a numeric or boolean

# fit the survival model
os_surv <- survfit(
    Surv(Months_of_follow_up_from_initial_diagnosis, Alive_with_last_f_up) ~ BRCA_EGFR_State,
    type="kaplan-meier",
    conf.type="log",
    data=surv_dat)

summary(os_surv)

survdiff(Surv(Months_of_follow_up_from_initial_diagnosis, Alive_with_last_f_up) ~ BRCA_EGFR_State, data=surv_dat)

plot(os_surv,
     col = seq_along(levels(surv_dat[["BRCA_EGFR_State"]])),
     lwd = 2,
     xlab="Months of Follow-up",
     ylab="Survival",
     main = "Overall Survival")
legend(x="topright",
       col=seq_along(levels(surv_dat[["BRCA_EGFR_State"]])),
       lwd=2,
       legend=levels(surv_dat[["BRCA_EGFR_State"]])
       )
```


# Is BRCA mutation an independent prognostic marker?

```{r}
surv_dat_subset <- surv_dat[surv_dat[["Baseline_LDH_at_Treatment_Initiation"]] %in% c(0,1), ]

res.cox <- coxph(Surv(Months_of_follow_up_from_initial_diagnosis, Alive_with_last_f_up) ~ BRCA_EGFR_State + Age_at_Initiation + Gender + Stage + Baseline_LDH_at_Treatment_Initiation, 
                 data =  surv_dat_subset)
summary(res.cox)
```

# Session

```{r}
sessionInfo()
```