---
title: "DiffBind Heatmaps"
author: "Stephen Kelly"
date: "1/20/2017"
output: 
  html_document: 
    keep_md: yes
    toc: true
    toc_float: true
    toc_depth: 3    
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup & Workflow {.tabset .tabset-pills}

## Hide

## Show

### Get DiffBind Files

The DiffBind ChIP-Seq datasets for differentially bound regions between samples per histone marks will be found and linked


```{r, engine='bash', eval=FALSE, cache=TRUE}
project_dir="/ifs/home/kellys04/projects/SmithLab_ChIpSeq_2016-03-10"
diffheatmap_dir="${project_dir}/project_notes/diffbind_heatmaps"
cd "$diffheatmap_dir"

diffbind_dir="${diffheatmap_dir}/diffbind_sheets"
mkdir -p "$diffbind_dir"

diffbind_data_location="/ifs/home/kellys04/projects/SmithLab_ChIpSeq_2016-03-10/project_notes/results_dir/diffbind_DESeq2/"

diffbind_files="$(find "$diffbind_data_location" -type f -name "diff_bind.*" -name "*.csv" -exec readlink -f {} \;)"

for diff_file in $diffbind_files; do
(
echo "$diff_file"
hist_mark="$(basename "$(dirname "$diff_file")")"
echo "$hist_mark"
out_name="$(echo "$(basename "$diff_file")" | sed "s/diff_bind/$hist_mark/")"
echo "$out_name"

cd "$diffbind_dir"
ln -fs "$diff_file" "$out_name"
)
done
```

Create the heatmaps; this needs to be submitted to the HPC cluster since the diffbind sheets are quite large

```{r, engine='bash', eval=FALSE, cache=TRUE}
project_dir="/ifs/home/kellys04/projects/SmithLab_ChIpSeq_2016-03-10"
diffheatmap_dir="${project_dir}/project_notes/diffbind_heatmaps"
cd "$diffheatmap_dir"

heatmap_output_dir="${diffheatmap_dir}/heatmaps"
mkdir -p "$heatmap_output_dir"

diffbind_dir="${diffheatmap_dir}/diffbind_sheets"
diffbind_files="$(find "$diffbind_dir" -name "*.csv" | sort)"

# iterate over the DiffBind files
for diff_file in $diffbind_files; do
# parse file name for items
echo "$diff_file"
diff_file_basename="$(basename "$diff_file")"
mark_ID="${diff_file_basename%%.*}"
echo "$mark_ID"
output_plot_file="${heatmap_output_dir}/${diff_file_basename%%.p100*}_heatmap.png"
echo "$output_plot_file"

output_plot_logdir="${heatmap_output_dir}/${diff_file_basename%%.p100*}_logs"
mkdir -p "$output_plot_logdir"

# regex to use in R for the status labels, since they don't pass well to the heredoc
D_regex='.*\\.D\\..*'
R_regex='.*\\.R\\..*'


# submit bash heredoc qsub job
    qsub -wd $PWD -o :${output_plot_logdir}/ -e :${output_plot_logdir}/ -j y <<E0F
set -x
# start R heredoc
Rscript - "${diff_file}" "${output_plot_file}" "$D_regex" "$R_regex" "${output_plot_logdir}" <<'E0F2'

args <- commandArgs(TRUE)

cat("R is running...\n")
cat("Script args are:\n")
print(args)

cat("\n---------------\n")
cat("Script contents:")
cat("\n---------------\n")
# cat( readLines( args ) , sep = "\n" ) 

diff_sheet_file <- args[1]
plot_output <- args[2]
d_label_regex <- args[3]
r_label_regex <- args[4]
log_dir <- args[5]

cat("R session data file:\n")
cat(file.path(log_dir, "session.Rdata"))

save.image(file = file.path(log_dir, "session.Rdata"), compress = TRUE, safe = FALSE)

# load the DiffBind sheet
diff_df <- read.csv(file = diff_sheet_file, check.names = FALSE, stringsAsFactors = FALSE)

# filter for FDR adjusted p value < 0.02
diff_df <- diff_df[which(diff_df[["p.value"]] <= 0.02), ]

# only run if there are enough entries..
if(nrow(diff_df) >= 2){
    # keep only the columns with the sample values
    # remove_cols <- c("seqnames", "start", "end", "width", "strand", "Conc", "Conc_D", 
    #                  "Conc_R", "Fold", "p.value", "FDR", "feature", "external_gene_name", 
    #                  "gene_biotype", "start_position", "end_position", "insideFeature", 
    #                  "distancetoFeature", "shortestDistance", "fromOverlappingOrNearest")
    
    # cols_to_keep <- colnames(diff_df)[ ! colnames(diff_df) %in% remove_cols]
    
    # subset for only cols we want
    # diff_df <- diff_df[,cols_to_keep]
    
    # re order the cols
    # diff_df <- diff_df[order(colnames(diff_df))]
    
    d_cols <- sort(colnames(diff_df)[grep(pattern = '^([[:alpha:]]*[[:punct:]]D)[[:punct:]][[:alpha:]].*$', perl = TRUE, x = colnames(diff_df))])
    r_cols <- sort(colnames(diff_df)[grep(pattern = '^([[:alpha:]]*[[:punct:]]R)[[:punct:]][[:alpha:]].*$', perl = TRUE, x = colnames(diff_df))])
    # diff_df <- diff_df[c(d_cols, r_cols)]
    
    # log2 the whole thing
    # diff_df <- log2(diff_df)
    
    # colors: blue-white-red
    colors <- colorRampPalette( c("#010F57", "#010F57", "#FAFAFA", "#B21212", "#B21212") )(50)
    
    # column labels
    annotation_col <- data.frame(status=colnames(diff_df[c(d_cols, r_cols)]), stringsAsFactors=F)
    rownames(annotation_col) <- colnames(diff_df[c(d_cols, r_cols)])
    annotation_col['status'] <- gsub(d_label_regex, "D", x=annotation_col[["status"]])
    annotation_col['status'] <- gsub(r_label_regex, "R", x=annotation_col[["status"]])
    
    # column labels colors
    ann_colors <- list(status = c("seagreen4", "orchid4"))
    names(ann_colors[["status"]]) <- unique(annotation_col[["status"]])
    
    
    library(pheatmap)
    
    # heatmap with clustering
    heatmap_with_clustering <- pheatmap(log2(diff_df[c(d_cols, r_cols)]), 
             color=colors, 
             border_color=NA, 
             scale="row", 
             cluster_rows=T, 
             cluster_cols=T, 
             clustering_distance_rows="euclidean", 
             clustering_distance_cols="euclidean", 
             clustering_method="complete", 
             fontsize_row=4, 
             fontsize_col=10, 
             show_rownames=F, 
             annotation_col=annotation_col, 
             annotation_colors=ann_colors, 
             width=12, height=8,
             filename=file.path(plot_output))
             
             write.table(x = cbind(clustering_order = heatmap_with_clustering[['tree_row']][['order']], diff_df), 
            file = paste0(plot_output, ".tsv"), 
            quote = FALSE, sep = '\t', col.names = TRUE, row.names = FALSE)
             
    # without columns clustering
    heatmap_without_col_clustering <- pheatmap(log2(diff_df[c(d_cols, r_cols)]), 
             color=colors, 
             border_color=NA, 
             cluster_cols = FALSE,
             cluster_rows=TRUE, 
             clustering_distance_rows="euclidean", 
             scale="row", 
             fontsize_row=4, 
             fontsize_col=10, 
             show_rownames=F, 
             annotation_col=annotation_col, 
             annotation_colors=ann_colors, 
             width=12, height=8,
             filename=paste0(plot_output, "_no-col_clust.png"))
             
             write.table(x = cbind(clustering_order = heatmap_without_col_clustering[['tree_row']][['order']], diff_df), 
            file = paste0(plot_output, "_no-col_clust.tsv"), 
            quote = FALSE, sep = '\t', col.names = TRUE, row.names = FALSE)



} else {
    print("ERROR: NOT ENOUGH ENTRIES IN THE FILE")
}

save.image(file = file.path(log_dir, "session.Rdata"), compress = TRUE, safe = FALSE)

E0F2
E0F
done

```

```

# Heatmaps 

```{r, results='asis'}
# {.tabset .tabset-pills}

heatmap_dir <- "/ifs/home/kellys04/projects/SmithLab_ChIpSeq_2016-03-10/project_notes/diffbind_heatmaps/heatmaps/"
heatmap_files <- dir(path = heatmap_dir, pattern = "*.png", full.names = TRUE)

for(hfile in heatmap_files){
    cat(sprintf('## %s \n \n', basename(hfile)))
    cat(sprintf("![](%s) \n \n", hfile))
}

```




