---
title: "Total Reads Report"
author: "Stephen Kelly"
date: "6/28/2017"
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
# http://rmarkdown.rstudio.com/html_document_format.html
# {.tabset .tabset-fade .tabset-pills}
knitr::opts_chunk$set(echo = TRUE)
# install.packages(c("ggplot2", "reshape2", "data.table", "knitr", "DT"))
library("knitr")
library("ggplot2")
library("reshape2")
library("data.table")
library("DT")
# ~~~~~ CUSTOM FUNCTIONS ~~~~~ #
mycat <- function(text){
    # function for formatting text in the report
    cat(gsub(pattern = "\n", replacement = "  \n", x = text))
}
```

```{r load_summary_tables}
results_dirs <- "/ifs/data/molecpathlab/NGS580_WES-development/validation_plots/results_dirs"

# there are symlink loops that are going to mess up R file search; 
# need to use this:
# 
# /ifs/data/molecpathlab/NGS580_WES-development/validation_plots$ find -L results_dirs/ -name "summary-combined.wes.csv" 2>/dev/null
# summary_combined_files <- dir(results_dirs, recursive = TRUE, pattern = "summary-combined.wes.csv", full.names = TRUE)
# summary_combined_files <- system("find -L /ifs/data/molecpathlab/NGS580_WES-development/validation_plots/results_dirs -name 'summary-combined.wes.csv' 2>/dev/null", intern = TRUE)
# find -L /ifs/data/molecpathlab/NGS580_WES-development/validation_plots/results_dirs -name 'summary-combined.wes.csv' -exec readlink -f {} \; 2>/dev/null | sort -u | xargs | sed 's| |",\n"|g'

# files to load data from
summary_combined_files <- c("/ifs/data/molecpathlab/NGS580_WES/170512_NB501073_0009_AHF5H2BGX2/results_2017-05-14_14-43-33/summary-combined.wes.csv",
"/ifs/data/molecpathlab/NGS580_WES/170519_NB501073_0010_AHCLLMBGX2/results_2017-05-24_14-15-19/summary-combined.wes.csv",
"/ifs/data/molecpathlab/NGS580_WES/170526_NB501073_0011_AHCJTYBGX2/results_2017-05-31_11-18-08/summary-combined.wes.csv",
"/ifs/data/molecpathlab/NGS580_WES/170602_NB501073_0012_AHCKYCBGX2/results_2017-06-05_09-51-00/summary-combined.wes.csv",
"/ifs/data/molecpathlab/NGS580_WES/170609_NB501073_0013_AHF7K3BGX2/results_2017-06-12_18-23-30/summary-combined.wes.csv",
"/ifs/data/molecpathlab/NGS580_WES/170616_NB501073_0014_AH2VVMBGX3/results_2017-06-19_10-05-42/summary-combined.wes.csv",
"/ifs/data/molecpathlab/NGS580_WES/170623_NB501073_0015_AHY5Y3BGX2/results_2017-06-26_20-11-26/summary-combined.wes.csv",
"/ifs/data/molecpathlab/NGS580_WES/NS17-01-35154137/results_2017-05-23_17-20-30/summary-combined.wes.csv",
"/ifs/data/molecpathlab/NGS580_WES/NS17-02/results_2017-05-23_17-38-30/summary-combined.wes.csv",
"/ifs/data/molecpathlab/NGS580_WES/NS17-03/results_2017-05-23_16-58-16/summary-combined.wes.csv",
"/ifs/data/molecpathlab/NGS580_WES/NS17-05/results_2017-05-24_13-24-30/summary-combined.wes.csv"
)

# load all of them into a single dataframe
summary_combined_table <- data.frame()

for(i in summary_combined_files){
    run_ID <- basename(dirname(dirname(i)))
    df <- read.delim(file = i, sep = ',')
    df[["run"]] <- run_ID
    
    if(nrow(summary_combined_table) < 1){
        summary_combined_table <- df
    } else {
        summary_combined_table <- rbind(summary_combined_table, df)
    }
    
} 

# fix colname
setnames(x = summary_combined_table, old = c("X.SAMPLE", "MEAN.COVERAGE", "MEDIAN.COVERAGE", "X._bases_above_10", "X._bases_above_50", "X._bases_above_100", "X._bases_above_500"), new = c("sample", "mean coverage", "median coverage", "pcnt bases above 10", "pcnt bases above 50", "pcnt bases above 100", "pcnt bases above 500"))


write.table(x = summary_combined_table[, c("run", "sample")], file = "/ifs/data/molecpathlab/NGS580_WES-development/validation_plots/SampleSummary.tsv", quote = FALSE, sep = '\t', row.names = FALSE)

```


# Read Mapping

Number of DNA reads per run

```{r read_mapping}
# make a smaller table just for read mapping stats
mapping_table <- summary_combined_table[, c("sample", "run")]

mapping_table["mapped"] <- summary_combined_table[["MAPPED.READS"]]
mapping_table["deduplicated"] <- summary_combined_table[["DEDUPLICATED.READS"]]

mapping_table <- reshape2::melt(mapping_table, id.vars=c("sample", "run"), variable.name="type", value.name="reads")

# convert to millions
mapping_table[["reads"]] <- as.numeric(mapping_table[["reads"]]) / 1e6

mapping_table[["type"]] <- factor(x = mapping_table[["type"]], levels = sort(unique(as.character(mapping_table[["type"]]), decreasing = TRUE)))


mapping_per_sample_plot <- ggplot(data = mapping_table, aes(x = sample, y = reads, fill = type)) + geom_bar(stat="identity", position = "dodge") + ggtitle("Sample Read Mapping") + coord_flip() + ylab("reads (millions)")


mapping_per_run_plot <- ggplot(data = mapping_table, aes(x = run, y = reads, fill = type)) + geom_boxplot() + coord_flip() + ggtitle("Mapped Reads per Run Across Samples") + ylab("reads (millions)")

```

```{r, fig.height=24, fig.width=12}
# mapping_per_sample_plot
```

```{r, fig.height=12, fig.width=12}
mapping_per_run_plot
```

# Target Coverage

Coverage of target regions

```{r mean_median_read_coverage}
# make a smaller table of just read coverages
coverage_table <- summary_combined_table[, c("sample", "run", "mean coverage", "median coverage")]
setnames(x = coverage_table, old = c("mean coverage", "median coverage"), new = c("mean", "median"))
coverage_table <- reshape2::melt(coverage_table, id.vars=c("sample", "run"), variable.name="type", value.name="coverage")

coverage_plot <- ggplot(data = coverage_table, aes(x = run, y = coverage, fill = type)) + geom_boxplot() + coord_flip() + ggtitle("Coverage of Target Regions per Run Across Samples") 
```


```{r coverage_cutoffs}
# make a smaller table with the percent of samples at given cutoffs
coverage_cutoff_table <- summary_combined_table[, c("sample", "run", "pcnt bases above 10", "pcnt bases above 50", "pcnt bases above 100", "pcnt bases above 500")]
setnames(x = coverage_cutoff_table, old = c("pcnt bases above 10", "pcnt bases above 50", "pcnt bases above 100", "pcnt bases above 500"), new = c("10", "50", "100", "500"))
coverage_cutoff_table <- reshape2::melt(coverage_cutoff_table, id.vars=c("sample", "run"), variable.name="cutoff", value.name="pcnt")

coverage_cutoff_plot <- ggplot(data = coverage_cutoff_table[which(coverage_cutoff_table[["cutoff"]] %in% c("100", "500")), ], aes(x = run, y = pcnt, fill = cutoff)) + geom_boxplot() + coord_flip() + ggtitle("Percent of Bases Above Coverage Cutoff per Run Across Samples")

```

```{r low_coverage_regions}
# find the number of regions with <50 and 0 coverage per sample

NGS580_targets_file <- "/ifs/data/molecpathlab/NGS580_WES/NGS580_targets.bed"
NGS580_targets <- read.delim(NGS580_targets_file, header = FALSE)
num_target_regions <- nrow(NGS580_targets)

# find the files with the annotations if the low regions; do this in the terminal and get the file list here
# /ifs/data/molecpathlab/NGS580_WES-development/validation_plots
# $ find -L results_dirs/ -path "*/sns-wes-downstream-analysis/sns-wes-coverage-analysis/*" -name "regions_coverage_below_50_annotation.tsv" -exec readlink -f {} \; 2>/dev/null | sort -u | xargs | sed 's| |",\n"|g'
# $ find -L results_dirs/ -path "*/sns-wes-downstream-analysis/sns-wes-coverage-analysis/*" -name "regions_with_coverage_0_annotation.tsv" -exec readlink -f {} \; 2>/dev/null | sort -u | xargs | sed 's| |",\n"|g'

regions_coverage_below_50_files <- c("/ifs/data/molecpathlab/NGS580_WES/170512_NB501073_0009_AHF5H2BGX2/results_2017-05-14_14-43-33/sns-wes-coverage-analysis/170512_NB501073_0009_AHF5H2BGX2_results_2017-05-14_14-43-33_regions_coverage_below_50_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/170519_NB501073_0010_AHCLLMBGX2/results_2017-05-24_14-15-19/sns-wes-coverage-analysis/170519_NB501073_0010_AHCLLMBGX2_results_2017-05-24_14-15-19_regions_coverage_below_50_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/170526_NB501073_0011_AHCJTYBGX2/results_2017-05-31_11-18-08/sns-wes-coverage-analysis/170526_NB501073_0011_AHCJTYBGX2_results_2017-05-31_11-18-08_regions_coverage_below_50_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/170602_NB501073_0012_AHCKYCBGX2/results_2017-06-05_09-51-00/sns-wes-coverage-analysis/170602_NB501073_0012_AHCKYCBGX2_results_2017-06-05_09-51-00_regions_coverage_below_50_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/170609_NB501073_0013_AHF7K3BGX2/results_2017-06-12_18-23-30/sns-wes-coverage-analysis/170609_NB501073_0013_AHF7K3BGX2_results_2017-06-12_18-23-30_regions_coverage_below_50_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/170616_NB501073_0014_AH2VVMBGX3/results_2017-06-19_10-05-42/sns-wes-coverage-analysis/170616_NB501073_0014_AH2VVMBGX3_results_2017-06-19_10-05-42_regions_coverage_below_50_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/170623_NB501073_0015_AHY5Y3BGX2/results_2017-06-26_20-11-26/sns-wes-coverage-analysis/170623_NB501073_0015_AHY5Y3BGX2_results_2017-06-26_20-11-26_regions_coverage_below_50_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/NS17-01-35154137/results_2017-05-23_17-20-30/sns-wes-coverage-analysis/NS17-01-35154137_results_2017-05-23_17-20-30_regions_coverage_below_50_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/NS17-02/results_2017-05-23_17-38-30/sns-wes-coverage-analysis/NS17-02_results_2017-05-23_17-38-30_regions_coverage_below_50_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/NS17-03/results_2017-05-23_16-58-16/sns-wes-coverage-analysis/NS17-03_results_2017-05-23_16-58-16_regions_coverage_below_50_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/NS17-05/results_2017-05-24_13-24-30/sns-wes-coverage-analysis/NS17-05_results_2017-05-24_13-24-30_regions_coverage_below_50_annotation.tsv")


regions_coverage_0_files <- c("/ifs/data/molecpathlab/NGS580_WES/170512_NB501073_0009_AHF5H2BGX2/results_2017-05-14_14-43-33/sns-wes-coverage-analysis/170512_NB501073_0009_AHF5H2BGX2_results_2017-05-14_14-43-33_regions_with_coverage_0_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/170519_NB501073_0010_AHCLLMBGX2/results_2017-05-24_14-15-19/sns-wes-coverage-analysis/170519_NB501073_0010_AHCLLMBGX2_results_2017-05-24_14-15-19_regions_with_coverage_0_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/170526_NB501073_0011_AHCJTYBGX2/results_2017-05-31_11-18-08/sns-wes-coverage-analysis/170526_NB501073_0011_AHCJTYBGX2_results_2017-05-31_11-18-08_regions_with_coverage_0_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/170602_NB501073_0012_AHCKYCBGX2/results_2017-06-05_09-51-00/sns-wes-coverage-analysis/170602_NB501073_0012_AHCKYCBGX2_results_2017-06-05_09-51-00_regions_with_coverage_0_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/170609_NB501073_0013_AHF7K3BGX2/results_2017-06-12_18-23-30/sns-wes-coverage-analysis/170609_NB501073_0013_AHF7K3BGX2_results_2017-06-12_18-23-30_regions_with_coverage_0_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/170616_NB501073_0014_AH2VVMBGX3/results_2017-06-19_10-05-42/sns-wes-coverage-analysis/170616_NB501073_0014_AH2VVMBGX3_results_2017-06-19_10-05-42_regions_with_coverage_0_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/170623_NB501073_0015_AHY5Y3BGX2/results_2017-06-26_20-11-26/sns-wes-coverage-analysis/170623_NB501073_0015_AHY5Y3BGX2_results_2017-06-26_20-11-26_regions_with_coverage_0_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/NS17-01-35154137/results_2017-05-23_17-20-30/sns-wes-coverage-analysis/NS17-01-35154137_results_2017-05-23_17-20-30_regions_with_coverage_0_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/NS17-02/results_2017-05-23_17-38-30/sns-wes-coverage-analysis/NS17-02_results_2017-05-23_17-38-30_regions_with_coverage_0_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/NS17-03/results_2017-05-23_16-58-16/sns-wes-coverage-analysis/NS17-03_results_2017-05-23_16-58-16_regions_with_coverage_0_annotation.tsv",
"/ifs/data/molecpathlab/NGS580_WES/NS17-05/results_2017-05-24_13-24-30/sns-wes-coverage-analysis/NS17-05_results_2017-05-24_13-24-30_regions_with_coverage_0_annotation.tsv")

run_names <- unique(c(basename(dirname(dirname(dirname(regions_coverage_below_50_files)))), basename(dirname(dirname(dirname(regions_coverage_0_files))))))

regions_coverage_below_50_files <- setNames(object = regions_coverage_below_50_files, nm = c(basename(dirname(dirname(dirname(regions_coverage_below_50_files))))))
regions_coverage_0_files <- setNames(object = regions_coverage_0_files, nm = basename(dirname(dirname(dirname(regions_coverage_0_files)))))

# load all of them into a single dataframe
summary_lowcoverage_table <- data.frame()

for(i in seq_along(regions_coverage_below_50_files)){
    run_ID <- names(regions_coverage_below_50_files[i])
    run_file <- regions_coverage_below_50_files[i]
    
    df <- read.delim(file = run_file)
    df[["run"]] <- run_ID
    df[["cutoff"]] <- "50"
    
    if(nrow(summary_lowcoverage_table) < 1){
        summary_lowcoverage_table <- df
    } else {
        summary_lowcoverage_table <- rbind(summary_lowcoverage_table, df)
    }
    
} 

for(i in seq_along(regions_coverage_0_files)){
    run_ID <- names(regions_coverage_0_files[i])
    run_file <- regions_coverage_0_files[i]
    
    df <- read.delim(file = run_file)
    df[["run"]] <- run_ID
    df[["cutoff"]] <- "0"
    
    if(nrow(summary_lowcoverage_table) < 1){
        summary_lowcoverage_table <- df
    } else {
        summary_lowcoverage_table <- rbind(summary_lowcoverage_table, df)
    }
    
} 

# figure out the percentage of regions with <50, 0 coverage
lowcoverage_aggre <- aggregate(feature ~ run + cutoff, data = summary_lowcoverage_table, FUN = length)
lowcoverage_plot <- ggplot(data = lowcoverage_aggre, aes(x = run, y = (feature / num_target_regions) * 100, fill = cutoff)) + geom_bar(stat = "identity", position = "dodge") + coord_flip() + ggtitle("Percent of Target Features at or below Coverage Cutoff per Run Across Samples") + ylab("Percent of features")
```


```{r, fig.height=12, fig.width=12}
coverage_plot
coverage_cutoff_plot
lowcoverage_plot
```

# Load worksheets

Load the cleaned up worksheet table with the IDs from the runs

```{r}
library("ggplot2")
library("reshape2")
library("data.table")

# validation_samples_table_file <- "/ifs/data/molecpathlab/NGS580_WES-development/validation_plots/Validation_Samples_Table.txt"
validation_samples_table_file <- "/ifs/data/molecpathlab/NGS580_WES-development/validation_plots/samples_index.txt"
validation_samples_table <- read.delim(file = validation_samples_table_file, header = TRUE, sep = '\t', stringsAsFactors = FALSE)
```


```{r map_sample_files}
# findthe path to the files per sample, add to the df
validation_samples_table[["summary_combined_file"]] <- NA
validation_samples_table[["regions_coverage_0_file"]] <- NA
validation_samples_table[["regions_coverage_below_50_file"]] <- NA

# average_coverage_per_sample_table <- read.delim(file.path(sns_wes_coverage_analysis_dir, "average_coverage_per_sample.tsv"))
# average_coverage_per_region_table <- read.delim(file.path(sns_wes_coverage_analysis_dir, "average_coverage_per_region.tsv"), header = FALSE)

for(i in seq(nrow(validation_samples_table))){
    # get items
    sample_id <- as.character(validation_samples_table[i, "sample"])
    results_path <- as.character(validation_samples_table[i, "NGS580_results_path"])
    coverage_dir <- file.path(results_path, "sns-wes-coverage-analysis")
    summary_combined_file <- dir(path = results_path, pattern = "summary-combined.wes.csv", full.names = TRUE)
    regions_coverage_0_file <- dir(path = coverage_dir, pattern = "_regions_with_coverage_0_annotation.tsv", full.names = TRUE)
    regions_coverage_below_50_file <- dir(path = coverage_dir, pattern = "_regions_coverage_below_50_annotation.tsv", full.names = TRUE)
    average_coverage_per_sample_file <- dir(path = coverage_dir, pattern = "_average_coverage_per_sample.tsv", full.names = TRUE)
    average_coverage_per_region <- dir(path = coverage_dir, pattern = "_average_coverage_per_region.tsv", full.names = TRUE)
    # sns-wes-coverage-analysis/NS17-05_results_2017-05-24_13-24-30_regions_with_coverage_0_annotation.tsv
    
    # add to df
    validation_samples_table[i, "summary_combined_file"] <- summary_combined_file
    validation_samples_table[i, "regions_coverage_0_file"] <- regions_coverage_0_file
    validation_samples_table[i, "regions_coverage_below_50_file"] <- regions_coverage_below_50_file
    validation_samples_table[i, "average_coverage_per_sample_file"] <- average_coverage_per_sample_file
    validation_samples_table[i, "average_coverage_per_region"] <- average_coverage_per_region
}

```

# Total reads by sample type

```{r combine_tables}
# get the files to pull data from
summary_combined_files <- unique(as.character(validation_samples_table[["summary_combined_file"]]))

# empty df to hold reformatted data
summary_df <- data.frame()

# get the data from each file, save it to df
for(i in seq_along(summary_combined_files)){
    
    sample_summary_combined_file <- summary_combined_files[i]
    run_ID <- basename(dirname(dirname(sample_summary_combined_file)))
    
    df <- read.delim(file = sample_summary_combined_file, sep = ',')
    
    df[["run"]] <- run_ID
    
    # fix the colnames
    setnames(x = df, old = c("X.SAMPLE", "MEAN.COVERAGE", "MEDIAN.COVERAGE", "X._bases_above_10", "X._bases_above_50", "X._bases_above_100", "X._bases_above_500"), new = c("sample", "mean coverage", "median coverage", "pcnt bases above 10", "pcnt bases above 50", "pcnt bases above 100", "pcnt bases above 500"))

    
    # add it to the full df
    if(nrow(summary_df) < 1){
        summary_df <- df
    } else {
        summary_df <- rbind(summary_df, df)
    }
    
} 


# nrow(summary_df) # 264

# merge the new data with the old data
summary_df <- merge(x = summary_df, 
                    y = validation_samples_table, 
                    by.x = c("sample", "run"), 
                    by.y = c("sample", "NGS580_RunID"))


# fix the empty tissue type entries
summary_df[which(summary_df[["tissue_type_IPMKB"]] == ""), "tissue_type_IPMKB"] <- "Unknown"
summary_df[which(summary_df[["sample_type"]] == ""), "sample_type"] <- "Unknown"
summary_df[which(summary_df[["is_cancer_normal"]] == ""), "is_cancer_normal"] <- "Unknown"
summary_df[which(summary_df[["pair_tumor_normal"]] == ""), "pair_tumor_normal"] <- NA
summary_df[which(summary_df[["pair_matched_sample"]] == ""), "pair_matched_sample"] <- NA

write.table(x = summary_df, file = '/ifs/data/molecpathlab/NGS580_WES-development/validation_plots/Validation_Sample_Summary_Table.tsv', quote = TRUE, sep = '\t', row.names = FALSE, col.names = TRUE)
```


Number of DNA reads per sample type


```{r read_mapping_per_type}
# make a smaller table just for read mapping stats

# make a copy of the df and start filtering out stuff we dont want in the plot
mapping_df <- data.frame()

# only include certain runs in the plot
# "NS17-03"  # Blood
# "NS17-13" # FFPE
# "NS17-09" # Fresh Frozen (only use NYU tumors)

mapping_df <- summary_df[ which(summary_df[["worksheet_NGS580_RunID"]] %in% c("NS17-13", "NS17-03")), ]
mapping_df <- rbind(mapping_df, subset(summary_df, subset = (sample_type == "Fresh_Frozen" & pair_tumor_normal == "Tumor") & grepl(pattern = '^NYU', x = sample) ))
mapping_df[which(mapping_df[["run"]] == "NS17-03" & mapping_df[["sample_type"]] == "Other"), "sample_type"] <- "Blood"

# reformat the data for plotting
cols_to_keep <- c("sample", "run", "sample_type", "worksheet_samples_in_the_run", "worksheet_NGS580_RunID", "is_paired_sample", "worksheet_Tissue_type")
# mapping_table <- mapping_df[, cols_to_keep]

mapping_df["total_reads"] <- mapping_df[["INPUT.READS"]]
# mapping_table["mapped"] <- mapping_df[["MAPPED.READS"]]
# mapping_table["deduplicated"] <- mapping_df[["DEDUPLICATED.READS"]]

# mapping_table <- reshape2::melt(mapping_table, id.vars = cols_to_keep, variable.name = "type", value.name = "reads")

# convert to millions of reads
mapping_df[["total_reads"]] <- as.numeric(mapping_df[["total_reads"]]) / 1e6

# re order the factor levels for the plot
# mapping_table[["type"]] <- factor(x = mapping_table[["type"]], levels = sort(unique(as.character(mapping_table[["type"]]), decreasing = TRUE)))
# mapping_per_sample_plot <- ggplot(data = mapping_table, aes(x = sample, y = reads, fill = sample_type)) + geom_bar(stat="identity", position = "dodge") + ggtitle("Sample Read Mapping") + coord_flip() + ylab("reads (millions)")

y_upper <- max(mapping_df[["total_reads"]]) * 1.05
mapping_per_type_plot <- ggplot(data = mapping_df, aes(x = sample_type, y = total_reads)) + geom_boxplot() + ggtitle("Reads per Sample Type") + ylab("reads (millions)") + scale_y_continuous(expand = c(0, 0), limits = c(0,y_upper))# + coord_flip()

```

## Reads plot

```{r, fig.height=8, fig.width=8}
mapping_per_type_plot
```

## Reads plot table {.tabset .tabset-pills}

### Hide

### Show

```{r}
datatable(mapping_df, 
          class = 'cell-border stripe', 
          rownames = FALSE, 
          caption = 'Runs and samples included in the plot',
          options = list(
              pageLength = 20
          )
)

```

-----

# Paired Sample Reads

Fresh Frozen samples from Paired Tumor/Normal runs:

170602_NB501073_0012_AHCKYCBGX2


```{r}
# paired_summary_df <- summary_df[which(summary_df[["run"]] %in% c("170602_NB501073_0012_AHCKYCBGX2", "170623_NB501073_0015_AHY5Y3BGX2")), ]
# paired_summary_df <- summary_df[which(summary_df[["sample_type"]] == "Fresh_Frozen"), ]
paired_summary_df <- summary_df[summary_df[["is_paired_sample"]] == "yes", ]
paired_summary_df <- subset(paired_summary_df,  grepl(pattern = '^NYU', x = sample) )

# reformat the data for plotting
# cols_to_keep <- c("sample", "run", "sample_type", "tissue_type_IPMKB","is_paired_sample", "pair_tumor_normal", "pair_matched_sample", "worksheet_samples_in_the_run", "worksheet_NGS580_RunID", "is_paired_sample", "worksheet_Tissue_type")
# paired_mapping_table <- paired_summary_df[, cols_to_keep]
# paired_mapping_table <- reshape2::melt(paired_mapping_table, id.vars = cols_to_keep, variable.name = "type", value.name = "reads")

paired_summary_df["total_reads"] <- paired_summary_df[["INPUT.READS"]]
# convert to millions of reads
paired_summary_df[["total_reads"]] <- as.numeric(paired_summary_df[["total_reads"]]) / 1e6

y_upper <- max(paired_summary_df[["total_reads"]]) + max(paired_summary_df[["total_reads"]])*0.05

pairs_mapping_per_run_plot <- ggplot(data = paired_summary_df, aes(x = pair_tumor_normal, y = total_reads)) + geom_boxplot() + ggtitle("Reads per Sample Type - NYGC Samples") + ylab("reads (millions)") + scale_y_continuous(expand = c(0, 0), limits = c(0,y_upper))# + facet_grid(~pair_tumor_normal) # + coord_flip()

```

```{r, fig.height=8, fig.width=8}
pairs_mapping_per_run_plot
```

## Paired Reads plot table {.tabset .tabset-pills}

### Hide

### Show

```{r}
datatable(paired_summary_df, 
          class = 'cell-border stripe', 
          rownames = FALSE, 
          caption = 'Runs and samples included in the plot',
          options = list(
              pageLength = 20
          )
)

```


# Reads per Number of Samples in the Run

Check the number of reads in runs that had 16 vs 24 samples


# All Reads by Sample by Run

```{r}

summary_df[["total_reads"]] <- as.numeric(summary_df[["INPUT.READS"]]) / 1e6
# summary_df$worksheet_NGS580_RunID
all_read_plot <- ggplot(data = summary_df, aes(x = sample, y = total_reads, fill = worksheet_NGS580_RunID)) + geom_bar(stat = "identity", position = "dodge") + coord_flip() + ggtitle("Reads per Sample") + ylab("reads (millions)") # group=interaction(sample_type, tissue_type_IPMKB)

pdf(file = "/ifs/data/molecpathlab/NGS580_WES-development/validation_plots/all_read.pdf", height = 50, width = 10)
print(all_read_plot)
dev.off()

```


```{r, fig.height=50, fig.width=10}
all_read_plot
```

# Low Coverage Regions for all runs

Check to make sure that the low coverage regions are the same across all runs

```{r}


```


# System Information {.tabset .tabset-pills}

## Hide

## Show


```{r}
# {.tabset .tabset-pills}

## Hide

## Show
# system info
mycat(sprintf("System:\n%s\n%s", system("hostname", intern = TRUE), system("uname -srv", intern = TRUE)))
mycat(sprintf("System user:\n%s", system("whoami", intern = TRUE)))
# dir
mycat(sprintf("System location:\n%s", system('pwd',intern=T)))

# repo info
mycat(sprintf("Git Remote:\n%s\n", system('git remote -v',intern=T)))
mycat(sprintf("Git branch and commit\n%s", system('printf "%s: %s" "$(git rev-parse --abbrev-ref HEAD)" "$(git rev-parse HEAD)"',intern=T)))

# date time
mycat(sprintf("Time and Date of report creation:\n%s", system("date", intern = TRUE)))

# R system info, packages, etc
sessionInfo()

# save current session
save.image(file="final_report_data.Rdata",compress = TRUE)
```
