---
title: "SeraCare SeraSeq Samples Analysis"
author: "Stephen Kelly"
date: "4/16/2018"
output:
  html_document:
    df_print: paged
    code_folding: hide
    fig_width: 10
    fig_height: 10

---
```{r setup, include=FALSE}
    # df_print: paged
knitr::opts_chunk$set(echo = TRUE)
library("ggplot2")
library("data.table")
library("binom")
library("plotly")
library("DT")
library("knitr")
background_seq_error_rate <- 0.02
power <- 0.95
alpha <- 0.05
conf_level <- 1 - alpha
```

Analysis of SeraCare SeraSeq synthetic DNA positive control target exome NGS580 panel analysis 

# Setup

- Load Data

```{r, cache=FALSE}

# file with paths to all SeraCare annotations files
annot_list_file <- "annot.txt"

# load the list of paths to files to be aggregated
annot_list_files <- readLines(annot_list_file)

# filter for LoFreq only 

annot_list_files <- annot_list_files[grepl(pattern = 'LoFreq', x = annot_list_files, ignore.case = TRUE)]

# add all entries to common df
annots <- data.frame()
for(annot_file in annot_list_files){
    # print(annot_file)
    # if (grepl(pattern = '180131_NB501073_0032_AHT5F3BGX3', x = annot_file)) break()
    # variant caller used
    type <- basename(dirname(annot_file)) # "VCF-LoFreq-annot"
    type <- gsub(pattern = 'VCF-', replacement = '', x = type)
    type <- gsub(pattern = '-annot', replacement = '', x = type) # LoFreq
    
    # results ID
    results <- basename(dirname(dirname(annot_file))) #  "results_2018-01-22_16-41-26"
    
    # run ID
    run <- basename(dirname(dirname(dirname(annot_file)))) # "180112_NB501073_0029_AHT5KFBGX3"
    
    # load file 
    df <- read.delim(file = annot_file, header = TRUE, sep = '\t', stringsAsFactors = FALSE, check.names = FALSE)
    
    # add extra fields
    sampleID <- unique(df[["SAMPLE"]]) # "SC-SERACARE"
    sample_run <- sprintf('%s-%s', sampleID, run) # "SC-SERACARE-180112_NB501073_0029_AHT5KFBGX3"
    df[["Run"]] <- run
    df[["Type"]] <- type
    df[["Results"]] <- results
    df[["Sample_Run"]] <- sample_run
    
    # merge against the full df
    cols1 <- colnames(annots)
    cols2 <- colnames(df)
    common_cols <- intersect(cols1, cols2)
    diff_cols <- setdiff(cols1, cols2)
    if(nrow(annots) < 1) {
        annots <- df
    } else {
        annots <- merge(x = annots, y = df, by = common_cols, all = TRUE)
    }
}

# fix some colnames
setnames(x = annots, old = c('#MUT'), new = c('MUT'))

# save a copy
write.table(x = annots, file = "all_SeraCare_annotations.tsv", quote = FALSE, sep = '\t', row.names = FALSE, col.names = TRUE)
saveRDS(object = annots, file = "all_SeraCare_annotations.Rds", compress = TRUE)
```

- Filter & Reformat

```{r, cache=FALSE}
# table with SeraCare SeraSeq mutations to filter the annotations against
selected_mutations_file <- "SeraCare_mutation_match_list.Sheet1.tsv"
selected_mutations <- read.delim(file = selected_mutations_file, header = TRUE, sep = '\t', stringsAsFactors = FALSE)

# filter for only variants with Amino Acid change that matches SC mutations
annots_match <- annots[annots[["AAChange.refGene"]] %in% selected_mutations[['NGS.580.LoFreq.Nomenclature']], ]

# merge the tables to get the SC table fields in the annotations table
annots_match_merge <- merge(x = annots_match, y = selected_mutations, by.x = "AAChange.refGene", by.y = 'NGS.580.LoFreq.Nomenclature', all.x = TRUE)

# clean colnames 
setnames(x = annots_match_merge, old = c("SeraCare.Ladder.Target.AF...."), new = c("True_AF"))

# add factor level for true AF value
annots_match_merge[["True_AF"]] <- factor(x = annots_match_merge[["True_AF"]], levels = sort(unique(annots_match_merge[["True_AF"]])))

# add numeric value of true AF
annots_match_merge[["True_AF_val"]] <- as.numeric(as.character(annots_match_merge[["True_AF"]])) / 100

# create an ID for each mutation
annots_match_merge[["MUT_ID"]] <- paste(annots_match_merge[["Gene.refGene"]], annots_match_merge[["SeraCare.HGVS.Nomenclature"]])



annots_match_merge[["coverage_required"]] <- apply(X = annots_match_merge, MARGIN = 1, FUN = function(row){
    coverage <- as.numeric(row["DEPTH"])
    AF <- as.numeric(row["FREQ"])
    True_AF <- as.numeric(row["True_AF_val"])
    
    coverage_required <- cloglog.sample.size(p.alt = True_AF, p = background_seq_error_rate, power = power, alpha = alpha)[["n"]]
    return(coverage_required)
})

annots_match_merge[["CI_lower"]] <- apply(X = annots_match_merge, MARGIN = 1, FUN = function(row){
    coverage <- as.numeric(row["DEPTH"])
    AF <- as.numeric(row["FREQ"])
    True_AF <- as.numeric(row["True_AF_val"])
    
    intervals <- binom.confint(x = True_AF * coverage, n = coverage, conf.level = conf_level, methods = "cloglog")
    CI_lower <- intervals[["lower"]]
    # CI_upper <- intervals[["upper"]]
    return(CI_lower)
})


annots_match_merge[["CI_upper"]] <- apply(X = annots_match_merge, MARGIN = 1, FUN = function(row){
    coverage <- as.numeric(row["DEPTH"])
    AF <- as.numeric(row["FREQ"])
    True_AF <- as.numeric(row["True_AF_val"])
    
    intervals <- binom.confint(x = True_AF * coverage, n = coverage, conf.level = conf_level, methods = "cloglog")
    # CI_lower <- intervals[["lower"]]
    CI_upper <- intervals[["upper"]]
    return(CI_upper)
})


# rearrange columns
annots_match_merge <- annots_match_merge[, c(colnames(annots_match_merge)[which(! colnames(annots_match_merge) %in% "AAChange.refGene")], "AAChange.refGene")]

# save a copy
write.table(x = annots_match_merge, file = "selected_SeraCare_annotations.tsv", quote = FALSE, sep = '\t', row.names = FALSE, col.names = TRUE)
saveRDS(object = annots_match_merge, file = "selected_SeraCare_annotations.Rds", compress = TRUE)



# function to apply more filters to the annotations
filter_df <- function(df, type = "LoFreq"){
    # filter for selected type of variant caller
    df <- df[which(df[["Type"]] == type), ]
    
    # filter out dilution samples
    df <- df[grep(pattern = "-1to4-", x = df[["SAMPLE"]], invert = TRUE), ]
    df <- df[grep(pattern = "-1to2-", x = df[["SAMPLE"]], invert = TRUE), ]
    df <- df[grep(pattern = "SeraCare-1to1", x = df[["SAMPLE"]], invert = TRUE), ]
    df <- df[grep(pattern = "250", x = df[["SAMPLE"]], invert = TRUE), ]
    df <- df[grep(pattern = "50ng", x = df[["SAMPLE"]], invert = TRUE), ]
    
    df <- df[grep(pattern = "171116_NB501073_0027_AHT5M2BGX3", x = df[["Run"]], invert = TRUE), ]
    df <- df[grep(pattern = "171004_NB501073_0026_AHL27YAFXX", x = df[["Run"]], invert = TRUE), ]
    df <- df[grep(pattern = "170918_NB501073_0025_AHH35JBGX3", x = df[["Run"]], invert = TRUE), ]
    df <- df[grep(pattern = "170809_NB501073_0019_AH5FFYBGX3", x = df[["Run"]], invert = TRUE), ]
    
    df <- df[grep(pattern = "SeraCare-1to1", x = df[["SAMPLE"]], invert = TRUE), ]
    df <- df[grep(pattern = "SC-2-SERACARE-180312_NB501073_0035_AH53C3BGX5", x = df[["Sample_Run"]], invert = TRUE), ]
    

    df <- droplevels(df)
    return(df)
}

# get just the matching LoFreq annotations
caller_type <- "LoFreq" # "GATK-HC" # "LoFreq" # "GATK-HC" # "LoFreq"
filtered_annot <- filter_df(annots_match_merge, type = caller_type)
# unique(filtered_annot[["Sample_Run"]])
# filtered_annot[["Run"]]

# replace all the Sample ID's in the table
filtered_annot[["Sample"]] <- factor(filtered_annot[["Sample_Run"]])
levels(filtered_annot[["Sample"]]) <- sprintf("SeraCare%s", seq(1, length(levels(filtered_annot[["Sample"]]))))

# calculate the standard deviations per Mutation
annot_sd <- aggregate(filtered_annot[["FREQ"]], by=list(as.factor(filtered_annot[["MUT_ID"]])), FUN=sd)
names(annot_sd) <- c("MUT_ID", "sd")

annot_mean <- aggregate(filtered_annot[["FREQ"]], by=list(as.factor(filtered_annot[["MUT_ID"]])), FUN=mean)
names(annot_mean) <- c("MUT_ID", "mean")

annot_mean_sd <- merge(x = annot_sd, y = annot_mean)
annot_mean_sd[["sd_interval_upper"]] <- annot_mean_sd[["mean"]] + ( 2 * annot_mean_sd[["sd"]] )
annot_mean_sd[["sd_interval_lower"]] <- annot_mean_sd[["mean"]] - ( 2 * annot_mean_sd[["sd"]] )

# merge the mean and SD values into the main dataframe
filtered_annot <- merge(x = filtered_annot, y = annot_mean_sd, by = "MUT_ID", all.x = TRUE)

# Add 'Variant' column
filtered_annot[['Variant']] <- filtered_annot[['MUT_ID']]

# save a copy
write.table(x = filtered_annot, file = sprintf("%s_filtered_SeraCare_annotations.tsv", caller_type), quote = FALSE, sep = '\t', row.names = FALSE, col.names = TRUE)
saveRDS(object = filtered_annot, file = sprintf("%s_filtered_SeraCare_annotations.Rds", caller_type), compress = TRUE)


```

- Make Plots

```{r, warning=FALSE, message=FALSE}
# Variant Allele Frequency Plot
vaf_plot <- ggplot(data =  filtered_annot, 
                   aes(y = FREQ, x = Sample, color = Variant, group = Variant, 
                       text = sprintf("True AF: %s\nQUAL: %s\nDEPTH: %s\nAA Change: %s", 
                                      True_AF, QUAL, DEPTH, SeraCar.Amino.Acid)) ) + 
    geom_point() +
    geom_line(alpha = 0.3) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    theme(panel.grid.minor = element_blank()) +
    scale_y_continuous(limits = c(0, NA), breaks = seq(0, max(max(filtered_annot[["CI_upper"]]), max(filtered_annot[["FREQ"]])), 0.025)) +
    ylab("Variant Allele Frequency") +
    ggtitle(sprintf('%s Detected Variant Frequencies', caller_type)) + 
    labs(color="Variant")
    

pdf(file = sprintf("%s_vaf_plot.pdf", caller_type), height = 10, width = 10)
print(vaf_plot)
invisible(dev.off())
vaf_plotly <- ggplotly(vaf_plot)
htmlwidgets::saveWidget(as_widget(vaf_plotly), file = sprintf("%s_vaf_plot.html", caller_type), selfcontained = TRUE)


# with 95% Confidence Intervals
vaf_CI_plot <- vaf_plot + 
    geom_linerange(aes(ymin = CI_lower, ymax = CI_upper)) +
    facet_grid(Variant~.) +
    theme(strip.text.y = element_text(angle = 0)) +
    theme(panel.grid.minor = element_blank()) +
    theme(legend.position="none")

pdf(file = sprintf("%s_vaf_CI_plot.pdf", caller_type), height = 40, width = 10)
print(vaf_CI_plot)
invisible(dev.off())


# with +/- 2std dev lines across per mut
vaf_SD_plot <- ggplot(data =  filtered_annot, 
       aes(y = FREQ, x = Sample, color = Variant, group = Variant) ) + 
    geom_point() +
    geom_line(alpha = 0.3) +
    geom_line(aes(y = mean), linetype="dotted") +
    geom_line(size = 1, aes(y = sd_interval_upper), alpha = 0.5) +
    geom_line(size = 1, aes(y = sd_interval_lower), alpha = 0.5) +
    facet_grid(Variant~.) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_y_continuous(limits = c(0, NA), 
                       breaks = seq(0, max(max(filtered_annot[["sd_interval_upper"]]), 
                                           max(filtered_annot[["FREQ"]])), 0.025)) +
    ylab("Variant Allele Frequency") +
    ggtitle(sprintf('%s Detected Variant Frequencies', caller_type)) +
    labs(color="Variant") +
    theme(strip.text.y = element_text(angle = 0)) +
    theme(panel.grid.minor = element_blank()) +
    theme(legend.position="none")

pdf(file = sprintf("%s_vaf_SD_plot.pdf", caller_type), height = 40, width = 10)
print(vaf_SD_plot)
invisible(dev.off())


# one plot per page
pdf(file = sprintf("all_vaf_CI_plots.pdf", caller_type), height = 8, width = 8)
for(mutation_ID in unique(filtered_annot[["Variant"]])){
    sub_df <- filtered_annot[ which(filtered_annot[["Variant"]] == mutation_ID), ]
    sub_vaf_plot <- ggplot(data =  sub_df, 
                           aes(y = FREQ, x = Sample, color = Variant, group = Variant, 
                               text = sprintf("True AF: %s\nQUAL: %s\nDEPTH: %s\nAA Change: %s", 
                                              True_AF, QUAL, DEPTH, SeraCar.Amino.Acid)) ) + 
        geom_point() +
        geom_line() + # alpha = 0.3
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1), panel.grid.minor = element_blank()) +
        scale_y_continuous(limits = c(0, NA), breaks = seq(0, max(max(sub_df[["CI_upper"]]), max(sub_df[["FREQ"]])) , 0.025)) +
        ylab("Variant Allele Frequency") +
        ggtitle(mutation_ID) + 
        labs(color="Variant") +
        geom_linerange(aes(ymin = CI_lower, ymax = CI_upper), alpha = 0.5, size = 1) +
        theme(strip.text.y = element_text(angle = 0)) +
        theme(panel.grid.minor = element_blank()) +
        theme(legend.position="none")
    print(sub_vaf_plot)
}
invisible(dev.off())

pdf(file = sprintf("all_vaf_SD_plots.pdf", caller_type), height = 8, width = 8)
for(mutation_ID in unique(filtered_annot[["Variant"]])){
    sub_df <- filtered_annot[ which(filtered_annot[["Variant"]] == mutation_ID), ]
    sub_vaf_plot <- ggplot(data =  sub_df, 
                           aes(y = FREQ, x = Sample, color = Variant, group = Variant, 
                               text = sprintf("True AF: %s\nQUAL: %s\nDEPTH: %s\nAA Change: %s", 
                                              True_AF, QUAL, DEPTH, SeraCar.Amino.Acid)) ) + 
        geom_point() +
        geom_line() + # alpha = 0.3
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1), panel.grid.minor = element_blank()) +
        scale_y_continuous(limits = c(0, NA), breaks = seq(0, max(max(sub_df[["CI_upper"]]), max(sub_df[["FREQ"]])) , 0.025)) +
        ylab("Variant Allele Frequency") +
        ggtitle(mutation_ID) + 
        labs(color="Variant") +
        geom_line(aes(y = mean), linetype="dotted") +
        geom_line(size = 1, aes(y = sd_interval_upper), alpha = 0.5) +
        geom_line(size = 1, aes(y = sd_interval_lower), alpha = 0.5) +
        theme(strip.text.y = element_text(angle = 0)) +
        theme(panel.grid.minor = element_blank()) +
        theme(legend.position="none")
    print(sub_vaf_plot)
}
invisible(dev.off())


# Depth of Coverage plot
cov_plot <- ggplot(data =  filtered_annot, 
                   aes(y = DEPTH, x = Sample, color = MUT_ID, group = MUT_ID) ) + 
    geom_point() +
    geom_line(alpha = 0.3) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_y_continuous(limits = c(0, NA)) +
    ylab("Depth of Coverage") +
    theme(panel.grid.minor = element_blank()) +
    ggtitle(sprintf('%s Variant Depth of Coverage', caller_type))

pdf(file = sprintf("%s_cov_plot.pdf", caller_type), height = 10, width = 10)
print(cov_plot)
invisible(dev.off())

# Quality Scores Plot
qual_plot <- ggplot(data =  filtered_annot, 
                   aes(y = QUAL, x = Sample, color = MUT_ID, group = MUT_ID) ) + 
    geom_point() +
    geom_line(alpha = 0.3) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_y_continuous(limits = c(0, NA)) +
    ylab("Quality Score") +
    theme(panel.grid.minor = element_blank()) +
    ggtitle(sprintf('%s Variant Quality Score', caller_type))

pdf(file = sprintf("%s_qual_plot.pdf", caller_type), height = 10, width = 10)
print(qual_plot)
invisible(dev.off())

```

# Samples

## Samples used in this analysis

```{r}
tmpdf <- filtered_annot[ which( ! duplicated(filtered_annot[c("SAMPLE", "Run")])), c("SAMPLE", "Sample", "Run")]
tmpdf <- tmpdf[order(tmpdf[["Run"]]), ]
rownames(tmpdf) <- NULL
kable(tmpdf, row.names = TRUE)
```

## All SeraCare SeraSeq samples

```{r}
tmpdf2 <- annots[ which( ! duplicated(annots[c("SAMPLE", "Run")])), c("SAMPLE", "Run")]
tmpdf2 <- tmpdf2[order(tmpdf2[["Run"]]), ]
rownames(tmpdf2) <- NULL
kable(tmpdf2, row.names = TRUE)
```

# Variants

```{r}
filtered_annot
```

# Variant Frequencies

```{r}
vaf_plot
```

## Frequencies with Standard Deviations

- dot: detected variant

- dotted horizontal line: group mean for given variant

- thick horizontal lines: 2 standard deviations above and below group mean

```{r, fig.height=40}
vaf_SD_plot
```

# Variant Depth of Coverage

```{r}
cov_plot
```

# Variant Quality Scores

```{r}
qual_plot
```

# Session

```{r}
sessionInfo()
save.image()
```
