---
title: "SeraSeq Dilution Variant Detection Limits"
author: "Stephen Kelly"
date: "8/29/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

Some positive control SeraCare samples were run on the NextSeq to determine the lower limit of detection for variants

Docs about the samples:
https://www.seracare.com/globalassets/resources/product-sheets/mkt-00335-rev-1-prod-sheet-seraseq-tri-level-tumor-mutation-dna-mix-v2-lc.pdf


NextSeq NGS580 analysis results dir:
```
/ifs/data/molecpathlab/NGS580_WES/170721_NB501073_0018_AH5C7GBGX3/results_2017-07-28_21-02-00
```

Need to check the variant calling results to find the quality and frequency of the variants

Also the samples were run at different
- amount of total sample DNA used (ng)
- dilution ratio of positive control sample SeraCare DNA with negative control HapMap DNA (to simulate mixed tumor cell samples)

# Setup

First, had to make a sample index in Excel for every sample and its dilution levels, and starting material amounts (`sample_index_file`)

Also had to manually create a table of the known mutations we want to look for in the analysis output (`selected_mutations_file`)

Load the data 
```{r}
sample_index_file <- "/ifs/data/molecpathlab/NGS580_WES-development/SeraSeq-dilution-detection-limit/sample-index.tsv"
sample_index <- read.delim(file = sample_index_file, header = TRUE, sep = '\t', stringsAsFactors = FALSE)
# > colnames(sample_index)
# [1] "sample"            "source"            "dilution"          "starting_material" "replicate"  
# > nrow(sample_index)
# [1] 24

# Already found the exact matching AAChange entries for every expected variant in the LoFreq output, saved in this file here
selected_mutations_file <- "/ifs/data/molecpathlab/NGS580_WES-development/SeraSeq-dilution-detection-limit/SeraCare_mutation_match_list.Sheet1.tsv"
selected_mutations <- read.delim(file = selected_mutations_file, header = TRUE, sep = '\t', stringsAsFactors = FALSE)
colnames(selected_mutations)
# [1] "Gene.ID"                       "SeraCare.HGVS.Nomenclature"    "SeraCar.Amino.Acid"            "Mutation.Type"                
# [5] "SeraCare.Ladder.Target.AF...." "NGS.580.LoFreq.Nomenclature"  
# colnames(selected_mutations) <- c("GeneIDs", "Coding", "Mutation_Type", "HGVS_Nomenclature", "AminoAcid", "TargetAF")

# variant calling output for GATK and LoFreq variant callers; results for all samples in a single table each
GATK_annotations_file <- "/ifs/data/molecpathlab/NGS580_WES/170721_NB501073_0018_AH5C7GBGX3/results_2017-07-28_21-02-00/sns-wes-downstream-analysis/170721_NB501073_0018_AH5C7GBGX3_results_2017-07-28_21-02-00_GATK_annotations.tsv"
loFreq_annotations_file <- "/ifs/data/molecpathlab/NGS580_WES/170721_NB501073_0018_AH5C7GBGX3/results_2017-07-28_21-02-00/sns-wes-downstream-analysis/170721_NB501073_0018_AH5C7GBGX3_results_2017-07-28_21-02-00_LoFreq_annotations.tsv"


GATK_annotations <- read.delim(file = GATK_annotations_file, header = TRUE, sep = '\t', stringsAsFactors = FALSE)
loFreq_annotations <- read.delim(file = loFreq_annotations_file, header = TRUE, sep = '\t', stringsAsFactors = FALSE)
# > colnames(loFreq_annotations)
#  [1] "X.MUT"              "SAMPLE"             "CHR"                "POS"                "QUAL"               "DEPTH"             
#  [7] "FREQ"               "Ref"                "Alt"                "Func.refGene"       "Gene.refGene"       "GeneDetail.refGene"
# [13] "ExonicFunc.refGene" "AAChange.refGene"   "snp138"             "snp138NonFlagged"   "ExAC_ALL"           "esp6500siv2_all"   
# [19] "X1000g2015aug_all"  "cosmic80"           "CADD13_PHRED"       "FATHMM_noncoding"   "FATHMM_coding"   

```

# Find the matches in the data

Need to search for known mutations from the SeraCare samples in the variant calling output tables. 
Thankfully this time I have the exact matches already listed in the `selected_mutations_file` file so I do not need to implement crazy pattern searching like last time to find the matching variant entries from the experimental output.

```{r}
lofreq_matches <- loFreq_annotations[loFreq_annotations[["AAChange.refGene"]] %in% selected_mutations[['NGS.580.LoFreq.Nomenclature']], ]
GATK_matches <- GATK_annotations[GATK_annotations[["AAChange.refGene"]] %in% selected_mutations[['NGS.580.LoFreq.Nomenclature']], ]
# use lofreq because it has higher sensitivity! 
```

Next we need to clean up the `lofreq_matches` dataframe some more for plotting.
- fix the "dilution" and "starting_material" column entries and convery them to ordered factors
- remove the HapMap-only samples from the df (they did not match any variants, which is to be expected)
- add a MutationID column
- save a hard copy of the table to file

```{r}
write.table(x = lofreq_matches, file = "/ifs/data/molecpathlab/NGS580_WES-development/SeraSeq-dilution-detection-limit/lofreq_matches.tsv", quote = FALSE, sep = '\t', row.names = FALSE, col.names = TRUE)

# get rid of the HapMap samples for the plots
lofreq_matches <- lofreq_matches[! grepl(pattern = "HapMap", x = lofreq_matches[["SAMPLE"]] ), ]
nrow(lofreq_matches)
# [1] 464

# add the sample_index fields, except for hapmap
sample_index_samples <- sample_index[! grepl(pattern = "HapMap", x = sample_index[["sample"]]), ]

# merge in the fields from the original tables
lofreq_matches <- merge(x = lofreq_matches, y = sample_index_samples, by.x = 'SAMPLE', by.y = 'sample', all = TRUE)
nrow(lofreq_matches)
# [1] 464

lofreq_matches <- merge(x = lofreq_matches, y = selected_mutations, by.x = 'AAChange.refGene', by.y = 'NGS.580.LoFreq.Nomenclature', all = TRUE)
nrow(lofreq_matches)
# [1] 464

# need to clean up the table some more; 
# - convert dilution, starting material to numerics

lofreq_matches[["dilution"]] <- gsub(pattern = "1|1", replacement = "1", x = lofreq_matches[["dilution"]], fixed = TRUE)
lofreq_matches[["dilution"]] <- gsub(pattern = "1|2", replacement = "0.5", x = lofreq_matches[["dilution"]], fixed = TRUE)
lofreq_matches[["dilution"]] <- gsub(pattern = "1|4", replacement = "0.25", x = lofreq_matches[["dilution"]], fixed = TRUE)

# reorder the factor levels
lofreq_matches[["dilution"]] <- factor(x = lofreq_matches[["dilution"]], levels = sort(as.numeric(unique(lofreq_matches[["dilution"]])), decreasing = TRUE) )

lofreq_matches[["starting_material"]] <- gsub(pattern = 'ng', replacement = '', x = lofreq_matches[["starting_material"]])
lofreq_matches[["starting_material"]] <- factor(x = lofreq_matches[["starting_material"]], levels = sort(as.numeric(unique(lofreq_matches[["starting_material"]])), decreasing = TRUE) )

# add unique ID columns for the mutations
lofreq_matches[["MutationID"]] <- paste(lofreq_matches[["Gene.ID"]], lofreq_matches[["SeraCar.Amino.Acid"]], sep = ' ')


# save hard copy
write.table(x = lofreq_matches, file = "/ifs/data/molecpathlab/NGS580_WES-development/SeraSeq-dilution-detection-limit/lofreq_matches_merged.tsv", quote = FALSE, sep = '\t', row.names = FALSE, col.names = TRUE)


# check that all original varaints are present in the final table
length(unique(selected_mutations[["NGS.580.LoFreq.Nomenclature"]])) # 35
length(unique(lofreq_matches[['AAChange.refGene']])) # [1] 35

any(! unique(selected_mutations[["NGS.580.LoFreq.Nomenclature"]]) %in% unique(lofreq_matches[['AAChange.refGene']]))
# [1] FALSE
```

# Plots

Now we will make some plots; we will start with `ggplot2` bar plots, then make them interactive to better explore the data with `plotly`. 

```{r}
library("ggplot2")
library("plotly")

# plot the depth 
mutation_plot <- ggplot(data = lofreq_matches, aes(x = MutationID, y = DEPTH, fill = starting_material)) +  
    facet_grid(starting_material~dilution) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    ggtitle("Variant Depth per Dilution Factor & Starting Material (ng)") +
    geom_bar(position = "dodge", stat = "identity") + 
    scale_y_continuous(breaks = seq(0, max(lofreq_matches[["DEPTH"]]), 100)) + 
    geom_hline(aes(yintercept = 200), linetype="dotted")

mutation_plotly <- ggplotly(mutation_plot, tooltip = "all")

pdf(file = '/ifs/data/molecpathlab/NGS580_WES-development/SeraSeq-dilution-detection-limit/lofreq_mutations_depth.pdf', width = 18, height = 14)
print(mutation_plot)
dev.off()

htmlwidgets::saveWidget(as_widget(mutation_plotly), file = "/ifs/data/molecpathlab/NGS580_WES-development/SeraSeq-dilution-detection-limit/lofreq_mutations_depth.html", selfcontained = TRUE)


# plot the freq
mutation_plot_freq <- ggplot(data = lofreq_matches, aes(x = MutationID, y = FREQ, fill = starting_material)) +  
    facet_grid(starting_material~dilution) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    ggtitle("Variant Frequency per Dilution Factor & Starting Material (ng)") +
    geom_bar(position = "dodge", stat = "identity") +
    scale_y_continuous(breaks = seq(0, max(lofreq_matches[["FREQ"]]), 0.01))

mutation_plotly_freq <- ggplotly(mutation_plot_freq, tooltip = "all")

pdf(file = '/ifs/data/molecpathlab/NGS580_WES-development/SeraSeq-dilution-detection-limit/lofreq_mutations_freq.pdf', width = 18, height = 14)
print(mutation_plot_freq)
dev.off()

htmlwidgets::saveWidget(as_widget(mutation_plotly_freq), file = "/ifs/data/molecpathlab/NGS580_WES-development/SeraSeq-dilution-detection-limit/lofreq_mutations_freq.html", selfcontained = TRUE)

```
